---
title: "H2A_Final"

We create cumulative micronutrient burden group variables and examine associations to telomere length at 28 months. 
The function "fit_RE_GLM" is used to model a GLM for each exposure and outcome pair. Covariate pre-screening and selection is built into the model code. 

The data was subseted to only the variables utilized for code optimization and reduce run-time. 
---
Downloading packages, getting data frame, and 0 function

```{r}
library(dplyr)
library(tidyr)
library(washbgam)
library(washb)
library(devtools)
library(here)
library(ggplot2)
library(here)
install.packages("car")
library(car)


setwd("/Users/sumuccu/Micronutrient-Telomere")


rm(list=ls())

source(here("0-config.R"))

d <- read.csv("/Users/sumuccu/Desktop/bangladesh-merged-eed-mn-data.csv")

```

Assembling cumulaitve burden score variables

We created four different burden scores

- All micronutrients 
- Non-Iron (ex. Vitamin A, vitamin B, and vitamin B12 ONLY)
- B Vitamins
- Vitamin A and Iron Deficiency Anemia 

We added a sensitivity analysis that utilized a missingness component (i.e. model will allow one missing domain). 

We also created a secondary analysis that compares 1 and 2+ micronutrient deficiencies to telomere length. 


```{r}

d <- d %>%
  select(lowfer, lowfer_inf, highsTFR, iron_def, irondef_inf , irondefanemia, irondefanemia_inf, vit_A_def, B12def, folatedef, delta_TS_Z,     
         logFER_inf, hemo_eeel, hepcidin_eeel,
         logsTFR_inf, logRBP_inf, logB12, logfolate, clusterid, TS_t3_Z, sex, age, 
         birthord, momage, momheight, momedu, hfiacat, Nlt18, Ncomp, watmin, walls, floor, HHwealth_scaled, tr, 
         life_viol_any_t3_cat, viol_any_preg_cat, sumscore_t0_mom_Z)

# Restrict H2A analyses to the telomere-available analytic sample.
d <- d %>%
  filter(!is.na(TS_t3_Z))

make_def_count <- function(dat, vars, min_nonmissing = length(vars)) {
  observed_n <- rowSums(!is.na(dat[, vars]))
  score <- rowSums(dat[, vars], na.rm = TRUE)
  score[observed_n < min_nonmissing] <- NA_real_
  score
}

summarize_burden_levels <- function(dat, score_var, max_domains, label, outcome_var = "TS_t3_Z") {
  score_sym <- rlang::sym(score_var)
  outcome_sym <- rlang::sym(outcome_var)

  dat %>%
    filter(!is.na(!!score_sym)) %>%
    count(level = !!score_sym, name = "n") %>%
    complete(level = 0:max_domains, fill = list(n = 0)) %>%
    mutate(
      composite = label,
      total_n = sum(n),
      percent = 100 * n / total_n
    ) %>%
    rowwise() %>%
    mutate(
      level_label = paste0(level, ifelse(level == 1, " deficiency", " deficiencies")),
      prop_ci = list(stats::binom.test(n, total_n)$conf.int),
      percent_lb = 100 * prop_ci[[1]][1],
      percent_ub = 100 * prop_ci[[1]][2],
      n_with_outcome = sum(!is.na(dat[[score_var]]) & dat[[score_var]] == level & !is.na(dat[[outcome_var]]))
    ) %>%
    ungroup() %>%
    select(composite, level, level_label, n, total_n, percent, percent_lb, percent_ub, n_with_outcome)
}

# Primary burden score:
# one non-overlapping micronutrient deficiency count per domain
# iron is represented once using the inflammation-adjusted iron deficiency indicator
d$mm_domain_primary <- make_def_count(
  d,
  vars = c("irondef_inf", "vit_A_def", "B12def", "folatedef")
)

# Sensitivity:
# same primary burden score, but allow one missing domain
d$mm_domain_primary_relaxed <- make_def_count(
  d,
  vars = c("irondef_inf", "vit_A_def", "B12def", "folatedef"),
  min_nonmissing = 3
)

# Sensitivity 1:
# non-iron burden only
d$mm_noiron <- make_def_count(
  d,
  vars = c("vit_A_def", "B12def", "folatedef")
)

# Sensitivity 2:
# B-vitamin burden only
d$mm_bdef <- make_def_count(
  d,
  vars = c("B12def", "folatedef")
)

# Sensitivity 3:
# focused anemia and vitamin A burden
d$mm_aanemia <- make_def_count(
  d,
  vars = c("vit_A_def", "irondefanemia_inf")
)

# Descriptive categorical version of the primary score
d$mm_domain_primary_cat <- case_when(
  is.na(d$mm_domain_primary) ~ NA_character_,
  d$mm_domain_primary == 0 ~ "0 deficiencies",
  d$mm_domain_primary == 1 ~ "1 deficiency",
  d$mm_domain_primary >= 2 ~ "2+ deficiencies"
)

d$mm_domain_primary_cat <- factor(
  d$mm_domain_primary_cat,
  levels = c("0 deficiencies", "1 deficiency", "2+ deficiencies")
)

fit_categorical_glm <- function(d, Y, cat_var, W = NULL) {
  contrast_specs <- list(
    list(
      new_var = "mm_primary_1_vs_0",
      label = "1 deficiency",
      keep_levels = c("0 deficiencies", "1 deficiency")
    ),
    list(
      new_var = "mm_primary_2plus_vs_0",
      label = "2+ deficiencies",
      keep_levels = c("0 deficiencies", "2+ deficiencies")
    )
  )

  contrast_res <- lapply(contrast_specs, function(spec) {
    dat_sub <- d %>%
      filter(.data[[cat_var]] %in% spec$keep_levels | is.na(.data[[cat_var]])) %>%
      mutate(
        !!spec$new_var := case_when(
          .data[[cat_var]] == spec$keep_levels[1] ~ 0,
          .data[[cat_var]] == spec$keep_levels[2] ~ 1,
          TRUE ~ NA_real_
        )
      )

    fit_res <- fit_washb_glm(
      d = dat_sub,
      Y = Y,
      X = spec$new_var,
      W = W,
      print = FALSE
    )

    fit_res %>%
      mutate(
        category = spec$label,
        Y = Y
      ) %>%
      select(category, Y, N, coef, se, zval, pval, lb, ub)
  })

  bind_rows(contrast_res)
}

```

running models

```{r}
Xvars <- c("mm_domain_primary", "mm_domain_primary_relaxed", "mm_noiron", "mm_bdef", "mm_aanemia")       

Yvars <- c("TS_t3_Z")

#Set list of adjustment variables

Wvars <- c(
  "sex", "age", "birthord", "momage", "momheight", "momedu",
  "hfiacat", "Nlt18", "Ncomp", "watmin", "walls", "floor", "HHwealth_scaled",
  "tr", "life_viol_any_t3_cat", "viol_any_preg_cat")

figure_names_h2a <- c(
  mm_domain_primary = "Micronutrient domain burden score",
  mm_domain_primary_relaxed = "Missing-data sensitivity score",
  mm_noiron = "Non-iron deficiency burden",
  mm_bdef = "B-vitamin deficiency burden",
  mm_aanemia = "Vitamin A and iron deficiency anemia burden",
  TS_t3_Z = "Telomere length z-score"
)

res_unadj <- NULL
for(i in Xvars){
  for(j in Yvars){
    print(i)
    temp <- fit_washb_glm(d=d, Y=j, X=i, W=NULL)
    temp$X <- i
    temp$Y <- j
    print(temp)
    res_unadj <- bind_rows(res_unadj, temp)
  }}

res_adj <- NULL
for(i in Xvars){
  for(j in Yvars){
    print(i)
    temp <- fit_washb_glm(d=d, Y=j, X=i, W=Wvars)
    temp$X <- i
    temp$Y <- j
    print(temp)
    res_adj <- bind_rows(res_adj, temp)
  }}

res_unadj_plot <- res_unadj %>%
  mutate(
    X = ifelse(X %in% names(figure_names_h2a), figure_names_h2a[X], X),
    Y = ifelse(Y %in% names(figure_names_h2a), figure_names_h2a[Y], Y)
  )

res_adj_plot <- res_adj %>%
  mutate(
    X = ifelse(X %in% names(figure_names_h2a), figure_names_h2a[X], X),
    Y = ifelse(Y %in% names(figure_names_h2a), figure_names_h2a[Y], Y)
  )

res_adj_main_plot <- res_adj_plot %>%
  filter(X != figure_names_h2a["mm_domain_primary_relaxed"])

p_h2a_main_domains <- ggplot(res_adj_main_plot, (aes(x=X, y=coef))) +
  geom_point(size=3, color = analysis_palette["primary"]) +
  geom_errorbar(aes(ymin=lb, ymax=ub), width = 0.75, linewidth = 1, color = analysis_palette["primary"]) +
  geom_hline(yintercept = 0, color = analysis_palette["neutral"], linetype = "dashed") +
  facet_wrap(~Y, ncol=2, scales="free") +
  coord_flip() +
  labs(
    title = "Main analysis",
    y = "Adjusted mean difference in telomere length z-score",
    x = NULL
  ) +
  theme_analysis_plot() +
  theme(axis.ticks.x=element_blank(),
        legend.position = "none",
        strip.text = element_text(vjust=1),
        axis.text.y=ggtext::element_markdown(),
        panel.spacing = unit(1, "lines"))

primary_cat_table <- d %>%
  filter(!is.na(mm_domain_primary_cat), !is.na(TS_t3_Z)) %>%
  group_by(mm_domain_primary_cat) %>%
  summarise(
    n = n(),
    mean_telomere_z = mean(TS_t3_Z, na.rm = TRUE),
    sd_telomere_z = sd(TS_t3_Z, na.rm = TRUE),
    se_telomere_z = sd_telomere_z / sqrt(n),
    mean_lb = mean_telomere_z - 1.96 * se_telomere_z,
    mean_ub = mean_telomere_z + 1.96 * se_telomere_z,
    .groups = "drop"
  )

burden_level_table <- bind_rows(
  summarize_burden_levels(
    dat = d,
    score_var = "mm_domain_primary",
    max_domains = 4,
    label = "Micronutrient domain burden score"
  ),
  summarize_burden_levels(
    dat = d,
    score_var = "mm_domain_primary_relaxed",
    max_domains = 4,
    label = "Missing-data sensitivity score"
  ),
  summarize_burden_levels(
    dat = d,
    score_var = "mm_noiron",
    max_domains = 3,
    label = "Non-iron deficiency burden"
  ),
  summarize_burden_levels(
    dat = d,
    score_var = "mm_bdef",
    max_domains = 2,
    label = "B-vitamin deficiency burden"
  ),
  summarize_burden_levels(
    dat = d,
    score_var = "mm_aanemia",
    max_domains = 2,
    label = "Vitamin A and iron deficiency anemia burden"
  )
)

burden_level_plot_data <- burden_level_table %>%
  mutate(
    composite = factor(
      composite,
      levels = c(
        "Micronutrient domain burden score",
        "Missing-data sensitivity score",
        "Non-iron deficiency burden",
        "B-vitamin deficiency burden",
        "Vitamin A and iron deficiency anemia burden"
      )
    ),
    level_label = factor(
      level_label,
      levels = c(
        "0 deficiencies",
        "1 deficiency",
        "2 deficiencies",
        "3 deficiencies",
        "4 deficiencies"
      )
    )
  )

p_h2a_burden_levels <- ggplot(
  burden_level_plot_data,
  aes(x = level_label, y = percent)
) +
  geom_col(
    fill = analysis_palette["fill"],
    color = analysis_palette["primary"],
    linewidth = 0.6
  ) +
  geom_errorbar(
    aes(ymin = percent_lb, ymax = percent_ub),
    width = 0.18,
    linewidth = 0.8,
    color = analysis_palette["primary"]
  ) +
  facet_wrap(~composite, scales = "free_x", ncol = 2) +
  theme_analysis_plot() +
  labs(
    title = "Distribution of deficiency levels within each burden score",
    x = "Number of deficient micronutrient domains",
    y = "Children in category (%)"
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    strip.text = element_text(hjust = 0)
  )

for (score_name in unique(as.character(burden_level_plot_data$composite))) {
  plot_dat <- burden_level_plot_data %>%
    filter(as.character(composite) == score_name)

  p_single_burden <- ggplot(
    plot_dat,
    aes(x = level_label, y = percent)
  ) +
    geom_col(
      fill = analysis_palette["fill"],
      color = analysis_palette["primary"],
      linewidth = 0.6
    ) +
    geom_errorbar(
      aes(ymin = percent_lb, ymax = percent_ub),
      width = 0.18,
      linewidth = 0.8,
      color = analysis_palette["primary"]
    ) +
    theme_analysis_plot() +
    labs(
      title = score_name,
      x = "Number of deficient micronutrient domains",
      y = "Children in category (%)"
    ) +
    theme(
      legend.position = "none",
      axis.text.x = element_text(angle = 0, hjust = 0.5)
    )

  safe_name <- gsub("[^A-Za-z0-9]+", "_", score_name)
  safe_name <- gsub("^_|_$", "", safe_name)

  ggsave(
    here("results", paste0("H2A_burden_levels_", safe_name, ".png")),
    p_single_burden,
    width = 6,
    height = 4,
    dpi = 600
  )
}

primary_cat_res <- fit_categorical_glm(
  d = d,
  Y = "TS_t3_Z",
  cat_var = "mm_domain_primary_cat",
  W = Wvars
) %>%
  mutate(
    category = factor(category, levels = c("1 deficiency", "2+ deficiencies")),
    Y_label = figure_names_h2a["TS_t3_Z"]
  )

p_h2a_main <- ggplot(primary_cat_res, aes(x = category, y = coef)) +
  geom_errorbar(
    aes(ymin = lb, ymax = ub),
    width = 0.12,
    linewidth = 1,
    color = analysis_palette["primary"]
  ) +
  geom_point(
    size = 3.5,
    color = analysis_palette["primary"]
  ) +
  geom_hline(yintercept = 0, color = analysis_palette["neutral"], linetype = "dashed") +
  theme_analysis_plot() +
  labs(
    title = "Preliminary analysis",
    subtitle = "Reference: no deficient micronutrient domains",
    x = "Number of deficient micronutrient domains",
    y = "Adjusted mean difference in telomere length z-score"
  ) +
  theme(legend.position = "none")

primary_missingness_sensitivity <- res_adj %>%
  filter(X %in% c("mm_domain_primary", "mm_domain_primary_relaxed")) %>%
  mutate(
    analysis = c("Complete-domain primary score", "Relaxed missing-data primary score"),
    analysis = factor(
      analysis,
      levels = c("Complete-domain primary score", "Relaxed missing-data primary score")
    )
  ) %>%
  select(analysis, N, coef, se, zval, pval, lb, ub, X, Y)

p_h2a_missingness <- ggplot(primary_missingness_sensitivity, aes(x = analysis, y = coef)) +
  geom_point(size = 3, color = analysis_palette["secondary"]) +
  geom_errorbar(aes(ymin = lb, ymax = ub), width = 0.15, linewidth = 1, color = analysis_palette["secondary"]) +
  geom_hline(yintercept = 0, color = analysis_palette["neutral"], linetype = "dashed") +
  coord_flip() +
  theme_analysis_plot() +
  labs(
    title = "Sensitivity analysis",
    x = NULL,
    y = "Adjusted mean difference in telomere length z-score"
  ) +
  theme(legend.position = "none")



write.csv(res_unadj, here("results/H2Aunadj_results.csv"))
write.csv(res_adj, here("results/H2Aadj_results.csv"))
write.csv(primary_cat_table, here("results/H2A_primary_burden_categories.csv"), row.names = FALSE)
write.csv(primary_cat_res, here("results/H2A_primary_categorical_model.csv"), row.names = FALSE)
write.csv(primary_missingness_sensitivity, here("results/H2A_missingness_sensitivity.csv"), row.names = FALSE)
write.csv(burden_level_table, here("results/H2A_burden_level_distribution.csv"), row.names = FALSE)


ggsave(here("results/H2A_main_categorical_burden_plot.png"), p_h2a_main, width=6, height=5, dpi=600)
ggsave(here("results/H2A_sensitivity_panel.png"), p_h2a_main_domains, width=7, height=5, dpi=600)
ggsave(here("results/H2A_missingness_sensitivity_plot.png"), p_h2a_missingness, width=6, height=4, dpi=600)
ggsave(here("results/H2A_burden_level_distribution_plot.png"), p_h2a_burden_levels, width=10, height=7, dpi=600)

primary_cat_table
primary_cat_res
primary_missingness_sensitivity
burden_level_table
p_h2a_burden_levels

```
